---
title: "CMAR Measurements"
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width = 8)

library(dplyr)
library(here)
library(lubridate)
library(plotly)

source(here("pages/import_data_and_params.R"))

dat_depth_raw <- readRDS(here("pages/data/depth_rolling_sd_reprocessed.rds")) %>%
  select(
    -c(int_sample, n_sample, rolling_sd_flag_sensor_depth_measured_m)
  ) 

dat_depth <- dat_depth_raw %>%
  ss_pivot_longer() %>%
  filter(
    # sensor read outside range
    !(station == "Olding Island" &
        deployment_range == "2022-Jun-03 to 2022-Sep-30" &
        depth_log == 3),
    # freshwater stations
    !(station %in% c("Piper Lake", "Hourglass Lake", "0193", "Sissiboo")),
    # suspect range
    !(station == "Long Beach" &
        deployment_range == "2020-Jul-16 to 2020-Nov-29" & depth_log == 5),
    !(station == "Sandy Cove" &
        deployment_range == "2020-Jul-16 to 2020-Nov-30" & depth_log == 5),
    !(station == "Tickle Island 1" &
        deployment_range == "2020-Oct-21 to 2021-Aug-25" & depth_log == 5)
  )

beaver <- readRDS(here("pages/data-reprocessed/beaver_point_2019-10-20.rds")) %>% 
  filter(
    timestamp_utc >= as_datetime("2020-01-01"),
    timestamp_utc < as_datetime("2020-02-01")
  ) %>% 
  select(
    sensor_type, sensor_serial_number, 
    timestamp_utc, sensor_depth_at_low_tide_m, sensor_depth_measured_m)

```


This page provides an overview of where and how sensor depth is measured through the CMAR Coastal Monitoring Program.

These figures and values include data collected up to the end of the 2022 field season and subjected to preliminary Quality Control.

# Locations

CMAR has collected depth data from `r length(na.omit(unique(dat_depth$station)))` stations in `r length(na.omit(unique(dat_depth$county)))` counties (@fig-depth-station-locations). 

```{r, fig-temp-station-units}
#| label: fig-depth-station-locations
#| fig-height: 7
#| fig-cap: Approximate location of stations with temperature data. 

# interactive map
st_locations %>% 
  inner_join(
    dat_depth %>% 
      select(COUNTY = county, STATION = station) %>% 
      distinct(), by = join_by(COUNTY, STATION)
  ) %>% 
  leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  add_circle_markers_county(county_pal = county_pal, size = 6) %>% 
  add_county_fill_legend(county_pal = county_pal)
  
```

# Number of Observations

```{r}
n_county_obs <- dat_depth_raw %>% 
  group_by(county) %>%  
  summarise(n = n()) %>% 
  mutate(
    n_percent = round(100 * n / sum(n), digits = 2),
    units = "m"
  ) %>% 
  arrange(desc(n_percent))

```

A large proportion of depth records are from `r n_county_obs[1, "county"]$county` County (`r n_county_obs[1, "n_percent"]$n_percent` %) and `r n_county_obs[2, "county"]$county` County (`r n_county_obs[2, "n_percent"]$n_percent` %), while a small proportion are from `r n_county_obs[nrow(n_county_obs), "county"]$county` (`r n_county_obs[nrow(n_county_obs), "n_percent"]$n_percent` %) and `r n_county_obs[nrow(n_county_obs)-1, "county"]$county` (`r n_county_obs[nrow(n_county_obs)-1, "n_percent"]$n_percent` %) Counties (@fig-depth-n-obs). 


```{r}
#| label: fig-depth-n-obs
#| fig-height: 8
#| fig-cap: The number of dissolved oxygen observations in each county. 

p <- plot_n_obs_county(n_county_obs)

ggplotly(p)
```

# Sensors

CMAR does not deploy sensors specifically to measure depth; however, some of the sensors used to measure other variables also measure depth. Most VR2AR units and some aquaMeasure units (DOTs and SALs) provide depth measurements.

The aquaMeasure sensors are programmed to record depth every 10 minutes, and have a precision of 0.1 m. this typically results in a clear sinusoidal signal that would be expected from tidal changes (@fig-aquameasure).

```{r}
#| fig-height: 4
#| fig-cap: Measured depth observations from an aquaMeasure sensor deployed in January 2020 at Beaver Point.
#| label: fig-aquameasure
#| warning: false

p <- beaver %>% 
  filter(sensor_type == "aquameasure") %>% 
  ss_ggplot_variables() +
  facet_wrap(~sensor_type) +
  geom_point(col = "grey20", size = 0.5) +
  scale_y_continuous("sensor_depth_measured_m")
   
ggplotly(p)

```

In contrast, the VR2AR units only record depth every hour and have a resolution of 1 m. This results in very coarse depth measurements that don't clearly show a sinusoidal signal, and can result in apparent spikes (@fig-vemco). 

```{r}
#| fig-height: 4
#| fig-cap: Measured depth observations from a VR2AR sensor deployed in January 2020 at Beaver Point.
#| label: fig-vemco
#| warning: false

p <- beaver %>% 
  filter(sensor_type == "vr2ar") %>% 
  ss_ggplot_variables() +
  facet_wrap(~sensor_type) +
  geom_point(col = "grey20", size = 0.5)  +
  scale_y_continuous("sensor_depth_measured_m")
   
ggplotly(p)
```

# Applications

CMAR takes advantage of the opportunistic depth sampling to verify the estimated depth of sensors on the string in the Depth Crosscheck Test. Abrupt changes in measured depth can indicate that a sensor string was moved during or between deployments. Measured depth that increases over the deployment could indicate substantial biofouling causing the sensors to sink below the intended depth. Measured depth can also provide an estimate of the tidal range in an area.

