---
title: "Thresholds (mg/L)"
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width = 8)

library(here)
library(lubridate)
library(plotly)
library(stringr)
library(summaryplots)

source(here("pages/import_data_and_params.R"))
source(here("functions/filter_out_suspect_obs.R"))


st_locations <- st_locations %>%
  left_join(do_units, by = "STATION")

# correct and filter DO (mg/L) data
dat_do_mg_per_l <- readRDS(here("pages/data/do_mg_per_l_rolling_sd_prelim_qc.rds")) %>%
  select(
    -c(int_sample, n_sample, rolling_sd_flag_dissolved_oxygen_uncorrected_mg_per_l)
  ) %>%
  ss_pivot_longer() %>%
  select(
    COUNTY = county,
    STATION = station,
    DEPLOYMENT_PERIOD = deployment_range,
    TIMESTAMP = timestamp_utc,
    everything()) %>%
  mutate(VARIABLE = "Dissolved Oxygen", UNITS = "mg/L") %>%
  filter_out_suspect_obs() %>%
  select(-c(VARIABLE, UNITS)) %>%
  rename(
    county = COUNTY,
    station = STATION,
    deployment_range = DEPLOYMENT_PERIOD,
    timestamp_utc = TIMESTAMP
  ) %>%
  ss_pivot_wider() %>%
  mutate(month = month(timestamp_utc, label = TRUE))

colnames(dat_do_mg_per_l) <- str_remove_all(colnames(dat_do_mg_per_l), "value_")
```

Separate thresholds were calculated for dissolved oxygen measured in units of percent saturation and dissolved oxygen measured in units of mg per L. This page describes the thresholds for mg per L. Click [here](../pages/do_thresholds.qmd) for the thresholds for percent saturation.

QC test thresholds for dissolved oxygen (mg/L) were based on historical Coastal Monitoring Program data. Preliminary quality control (QC) was applied to the historical data. Obvious outliers, suspected biofouling, and freshwater stations were omitted.

To date, dissolved oxygen (mg/L) has only been measured in two counties[^1] in adjacent waterbodies[^2]. A single set of thresholds was determined for these counties.

The dissolved oxygen (mg/L) data has not been corrected for salinity. The threshold were calculated from and applied to the uncorrected data.

[^1]: Halifax and Lunenburg
[^2]: Mahone Bay and St. Margaret's Bay


# Gross Range Test

## Sensor Thresholds

The sensor thresholds were determined based on the manual for the HOBO DO (Table 1).

::: panel-tabset

### Table 1

Table 1: Dissolved oxygen (mg/L) sensor thresholds for the Gross Range Test.
```{r}
sensors %>%
  filter(str_detect(variable, "dissolved_oxygen"), sensor_type == "hobo do") %>%
  mutate(
    `Sensor (link to spec sheet)` =
      paste0('<a  target=_blank href=', url, '>', sensor_type,'</a>')
  ) %>%
  select(`Sensor (link to spec sheet)`, sensor_min, sensor_max) %>%
  datatable(options = dt_options2, rownames = FALSE, escape = FALSE)
```
:::

## User Thresholds

The dissolved oxygen (mg/L) observations are relatively normally distributed (@fig-do-hist), and so the mean and standard deviation was used to determine $user_{min}$ and $user_{max}$. The statistics and threshold values are shown in Table 2.

:::panel-tabset

### Figure 1
```{r}
#| warning: false
#| fig-height: 4
#| fig-cap: Distribution of dissolved oxygen (mg/L) observations (binwidth = 0.5 mg/L). Dotted orange lines indicate the user thresholds.
#| label: fig-do-hist

user_do <- dat_do_mg_per_l %>% 
  qc_calculate_user_thresholds(
    var = "dissolved_oxygen_uncorrected_mg_per_l",
    n_sd = 3, keep_stats = TRUE
  ) %>% 
  pivot_wider(names_from = threshold, values_from = threshold_value) %>% 
  select(mean = mean_var, stdev = sd_var, user_min, user_max) 

p <- plot_histogram(
  dat_do_mg_per_l, hist_col = dissolved_oxygen_uncorrected_mg_per_l, 
  binwidth = 0.5,
  pal= hist_pal, user_thresh = c(user_do$user_min, user_do$user_max)
) +
  scale_x_continuous("Dissolved Oxygen (mg/L)")

ggplotly(p, tooltip = "text")
```

### Table 2

Table 2: Gross Range Test statistics and user thresholds for dissolved oxygen (mg/L).
```{r}
user_do %>% 
  datatable(options = dt_options2, rownames = FALSE)
```
:::

# Climatological Test

@fig-do-clim shows the monthly mean and standard deviation of dissolved oxygen (mg/L) data. The observations are normally distributed within each month (@fig-do-hist-clim), and so the mean and standard deviation were used to calculate the seasonal thresholds (Table 3).

::: panel-tabset

### Figure 2
```{r}
#| fig-height: 5
#| fig-cap: Mean +/- 3 standard deviations of the monthly dissolved oxygen (mg/L) observations.
#| label: fig-do-clim
#| message: false

do_clim <- dat_do_mg_per_l %>% 
  group_by(month) %>% 
  summarise(
    mean = mean(dissolved_oxygen_uncorrected_mg_per_l), 
    stdev = sd(dissolved_oxygen_uncorrected_mg_per_l)
  ) %>% 
  mutate(
    variable = "dissolved_oxygen_uncorrected_mg_per_l",
    season_min = round(mean - 3 * stdev, digits = 3),
    season_max = round(mean + 3 * stdev, digits = 3)
  )

# # quick check
# do_thresh <- thresholds %>%
#   filter(variable == "dissolved_oxygen_uncorrected_mg_per_l", qc_test == "climatology") %>%
#   pivot_wider(names_from = threshold, values_from = threshold_value)
# 
# all.equal(do_clim$season_min, do_thresh$season_min)
# all.equal(do_clim$season_max, do_thresh$season_max)

p <- plot_mean_sd_season(do_clim, n_sd = 3, facet_county = FALSE) +
  scale_y_continuous("Dissolved Oxygen (mg/L)")
  
ggplotly(p)

```

### Figure 3
```{r}
#| warning: false
#| fig-height: 10
#| fig-cap:  Seasonal distribution of dissolved oxygen (mg/L) observations (binwidth = 0.5 mg/L). Dotted orange lines indicate the user thresholds. 
#| label: fig-do-hist-clim
#| message: false

p <- dat_do_mg_per_l %>% 
  plot_histogram(
    hist_col = dissolved_oxygen_uncorrected_mg_per_l, binwidth = 0.5,
    pal= hist_pal
  ) +
  geom_vline(
    data = do_clim, aes(xintercept = season_min), linetype = 3, col = "#EDA247"
  ) +
  geom_vline(
    data = do_clim, aes(xintercept = season_max), linetype = 3, col = "#EDA247"
  ) +
  scale_x_continuous("Dissolved Oxygen (mg/L)") +
  facet_wrap(~month,  ncol = 3) +
  theme(strip.background = element_rect(color = 1)) +
  theme_facet_plotly

ggplotly(p, tooltip = "text")
```

### Table 3

Table 2: Climatology statistics and thresholds for dissolved oxygen (mg/L).
```{r}
do_clim %>% 
  mutate( 
    season_mean = round(mean, digits = 3), 
    season_stdev = round(stdev, digits = 3)
  ) %>% 
  select(month, season_mean, season_stdev, season_min, season_max) %>% 
  datatable(options = dt_options, rownames = FALSE)
```

:::


# Spike Test

The distribution of the spike value is skewed right (@fig-do-spike-dist), and so several upper quartile values (90th, 95th, and 99.7th quartile) were evaluated to use as the $spike_{low}$. 

There were relatively few large single-value spikes in the dissolved oxygen (mg/L) data, and so the 99.7th quartile was selected to avoid false positives. $spike_{high}$ was defined as 3 * $spike_{low}$ to identify especially egregious spike values.

::: panel-tabset

### Figure 4
```{r}
#| warning: false
#| fig-height: 4
#| fig-cap: Distribution of the spike value of dissolved oxygen (percent saturation) observations (binwidth = 0.1 %). Dotted orange line indicates spike_low; dotted red line indicates spike_high.
#| label: fig-do-spike-dist

spike_thresh <- thresholds %>% 
  filter(variable == "dissolved_oxygen_uncorrected_mg_per_l", qc_test == "spike") %>% 
  pivot_wider(names_from = threshold, values_from = threshold_value)

dat_sp <- dat_do_mg_per_l %>% 
  group_by(county, station, deployment_range, sensor_serial_number) %>%
  dplyr::arrange(timestamp_utc, .by_group = TRUE) %>%
  mutate(
    lag_value = lag(dissolved_oxygen_uncorrected_mg_per_l),
    lead_value = lead(dissolved_oxygen_uncorrected_mg_per_l),
    spike_ref = (lag_value + lead_value) / 2,
    spike_value = abs(dissolved_oxygen_uncorrected_mg_per_l - spike_ref)  
  ) %>% 
  filter(!is.na(spike_value))

p <- dat_sp %>% 
  plot_histogram(
    hist_col = spike_value, binwidth = 0.1, pal = hist_pal,
    user_thresh = spike_thresh$spike_low
  ) +
  geom_vline(xintercept = spike_thresh$spike_high, col = "#DB4325", linetype = 3) +
  scale_x_continuous("Dissolved Oxygen (mg/L)\nSpike Value")

ggplotly(p, tooltip = "text")

```

### Table 4

Table 4: Spike thresholds for dissolved oxygen (mg/L).
```{r}
spike_thresh %>% 
  select(spike_low, spike_high) %>% 
  datatable(options = dt_options2, rownames = FALSE)
```

:::

# Rolling Standard Deviation Test

The distribution of rolling standard deviation is skewed right (@fig-do-roll-sd-dist), and so several upper quartile values were evaluated to use as the $rolling\_sd\_max$. 

The 90th, 95th, and 99.7th quartile values were each applied to the the raw data (no preliminary QC) and the results inspected. The 90th quartile was determined to be too stringent, as it generated false positives, while The 99.7th quartile was determined to be too lenient, as it resulted in false negatives. The 95th quartile identified 23 % of the raw data (no preliminary QC) as [Suspect/Of Interest]{style="color: #EDA247;"} and was considered the most useful threshold.

As discussed for dissolved oxygen (percent saturation), this is a very high percent of observations to flag as [Suspect/Of Interest]{style="color: #EDA247;"}. However, it was anticipated that a large proportion of the dissolved oxygen observations would be flagged by this test because there were substantial biofouling signals identified during the preliminary QC process. CMAR is currently assessing options to reduce biofouling on dissolved oxygen sensors.

There is no [Fail]{style="color: #DB4325;"} flag for the Rolling Standard Deviation Test because of the natural variability of the Water Quality variables. However, CMAR recommends that the dissolved oxygen observations flagged as [Suspect/Of Interest]{style="color: #EDA247;"} should be considered [Suspect]{style="color: #EDA247;"}, and filtered out of most analyses. 

::: panel-tabset

### Figure 5
```{r}
#| warning: false
#| fig-height: 4
#| fig-cap: Distribution of the 24-hour rolling standard deviation of dissolved oxygen (mg/L) observations (binwidth = 0.1 mg/L). Shaded green area shows observations that wil be flagged “Pass”, and shaded orange area shows observations that will be flagged as “Suspect / Of Interest" by the Rolling Standard Deviation Test.
#| label: fig-do-roll-sd-dist

do_roll_sd <- thresholds %>% 
  filter(
    variable == "dissolved_oxygen_uncorrected_mg_per_l",
         qc_test == "rolling_sd"
    ) %>% 
  select(threshold, threshold_value) %>% 
  pivot_wider(names_from = threshold, values_from = threshold_value) 

p <- dat_do_mg_per_l %>% 
  filter(!is.na(sd_roll)) %>% 
    plot_histogram(
    hist_col = sd_roll, binwidth = 0.1, pal = hist_pal,
    user_thresh = do_roll_sd$rolling_sd_max
  ) +
  scale_x_continuous("Dissolved Oxygen (mg/L)\nRolling Standard Deviation")

ggplotly(p, tooltip = "text")

```

### Table 5

Table 5: Rolling standard deviation threshold for dissolved oxygen (mg/L).
```{r}
do_roll_sd %>% 
  datatable(options = dt_options2, rownames = FALSE)
```
:::
