---
title: "CMAR Measurements"
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width = 8)

library(here)

source(here("pages/import_data_and_params.R"))

dat <- dat %>% filter(variable == "Temperature")
```

This page provides an overview of where and how dissolved oxygen is measured through the CMAR Coastal Monitoring Program.

These figures and values include data collected up to the end of the 2022 field season and subjected to preliminary Quality Control.

# Locations

CMAR has collected temperature data from `r length(na.omit(unique(dat$station)))` stations in `r length(na.omit(unique(dat$county)))` counties (@fig-temp-station-locations). 

```{r, fig-temp-station-units}
#| label: fig-temp-station-locations
#| fig-height: 7
#| fig-cap: Approximate location of stations with temperature data. 

leaflet(st_locations) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  add_circle_markers_county(county_pal = county_pal, size = 6) %>% 
  add_county_fill_legend(county_pal = county_pal)
  
```

# Number of Observations
```{r}
n_county_obs <- dat %>% 
  filter(group == "county") %>% 
  group_by(county) %>%  
  summarise(n = sum(n)) %>% 
  mutate(n_percent = round(100 * n / sum(n), digits = 2)) %>% 
  arrange(desc(n_percent))
```

A large proportion of these records are from `r n_county_obs[1, "county"]$county` County
(`r n_county_obs[1, "n_percent"]$n_percent` %), while a small proportion are from `r n_county_obs[nrow(n_county_obs), "county"]$county` (`r n_county_obs[nrow(n_county_obs), "n_percent"]$n_percent` %) and `r n_county_obs[nrow(n_county_obs)-1, "county"]$county` (`r n_county_obs[nrow(n_county_obs)-1, "n_percent"]$n_percent` %) Counties (@fig-temp-n-obs). 

```{r}
#| label: fig-temp-n-obs
#| fig-height: 8
#| fig-cap: The number of temperature observations in each county. 

p <- dat %>%
  filter(group == "county") %>% 
  plot_n_obs_county()

ggplotly(p)

```

# Depth

The placement of the temperature sensors depends on the depth of the water at each station. Typically, sensors are fastened 2, 5, 10, 15, and 20 m below the surface at low tide. At deeper stations, additional sensors are added every 5 - 10 m. At shallow stations, sensors maybe be attached nearer to the surface (@fig-temp-n-obs-depth).

Sensors may be placed at other depths for technical reasons or specific research projects (e.g., investigation of the oxycline in Whycocomagh Basin).

```{r}
#| label: fig-temp-n-obs-depth
#| fig-height: 10
#| fig-cap: Number of temperature observations at each depth (rounded to nearest whole number). 

p <- dat %>%
  filter(group == "all_depth") %>%
  group_by(depth) %>%
  summarise(n = sum(n)) %>%
  plot_n_obs_depth(text_size = 12, facet_county = FALSE)

ggplotly(p, tooltip = "text")
```

# Sensors

CMAR uses several types of sensors to measure temperature (Table 1):  

```{r}
sensors %>%
  select(-c(sensor_type, sensor_min, sensor_max)) %>% 
  mutate(
    variable = case_when(
      str_detect(variable, "temp") ~ "Temperature",
      str_detect(variable, "dissolved") ~ "Dissolved Oxygen",
      str_detect(variable, "salinity") ~ "Salinity",
      str_detect(variable, "depth") ~ "Pressure (sensor depth)"
    )
  ) %>% 
  pivot_wider(names_from = "variable", values_from = "variable") %>% 
  mutate(
    `Variable(s) Measured` = paste(
      Temperature, `Dissolved Oxygen`, Salinity, `Pressure (sensor depth)`,
      sep = ", "
    ),
    `Variable(s) Measured` = str_remove_all(`Variable(s) Measured`, "NA,|, NA"),
    `Sensor (link to manual)` = 
      paste0('<a  target=_blank href=', url, '>', sensor_model,'</a>')
  ) %>% 
  select(`Sensor (link to manual)`, Manufacturer = manufacturer, `Variable(s) Measured`) %>%
  arrange(Manufacturer) %>% 
  datatable(
    options = list(
      dom = 'ft',
      searching = FALSE,
      paging = FALSE,
      pageLength = 500,
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
    ), 
    rownames = FALSE, escape = FALSE,
    caption = "Table 1: Temperature sensors used in CMAR's Coastal Monitoring Program. Note that some aquaMeasure DOT, aquaMeausure SAL, and Vemco units do not measure pressure."
  )
```

<br>

Some of these sensors also measure other variables including dissolved oxygen, salinity, depth, and acoustic detections.

The Vemco VR2AR is typically the deepest sensor, anchored about 0.5 m above the sea floor. It has an acoustic release that is triggered to retrieve the sensor string. Deployments without VR2AR sensors are usually accessible from the surface, but some are retrieved by dragging or divers.




