---
title: "Thresholds"
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width = 8)

library(here)

source(here("pages/import_data_and_params.R"))

dat_filt <- dat_filt %>% filter(variable == "Temperature")
```


Most QC test thresholds for temperature were based on historical Coastal Monitoring Program data. Preliminary quality control was applied to this data[^1], leaving observations from `r length(na.omit(unique(dat_filt$station)))` stations in `r length(na.omit(unique(dat_filt$county)))` counties (@fig-temp-station-locations). 

[^1]: e.g., obvious outliers and freshwater stations were omitted

```{r}
#| label: fig-temp-station-locations
#| fig-height: 7
#| fig-cap: Approximate location of stations with temperature data used to inform QC test thresholds. Marker size is proportional to the number of temperature observations within the county.

# join the station locations dataset with the number of obs from each station
st_locations <- st_locations %>% 
  inner_join(
    dat_filt %>% 
      filter(group == "all_station", variable == "Temperature") %>% 
      select(COUNTY = county, STATION = station, n),
    by = c("COUNTY", "STATION") 
  ) %>% 
  group_by(COUNTY) %>% 
  mutate(
    n_tot = sum(n),
    n_prop = round(n / sum(n), digits = 2),
    popup = paste(COUNTY, STATION, n_prop, sep = "</br>")
  ) %>% 
  ungroup()

# interactive map
leaflet(st_locations) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  add_circle_markers_county(
    county_pal = county_pal, 
    size =  ~n_prop * 25
  ) %>% 
  add_county_fill_legend(county_pal = county_pal)
  
```

## Gross Range Test

### Sensor Thresholds

The sensor thresholds were determined based on the associated manual (Table 1).

```{r}
sensors %>% 
  filter(variable == "temperature_degree_c") %>%
  mutate(
    `Sensor Type (link to manual)` = 
      paste0('<a  target=_blank href=', url, '>', sensor_type,'</a>')
  ) %>% 
  select(`Sensor Type (link to manual)`, sensor_min, sensor_max ) %>%
  distinct() %>% 
  datatable(
    options <- list(
      dom = 'ft',
      searching = FALSE,
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
    ), 
    rownames = FALSE, escape = FALSE,
    caption = "Table 1: Temperature sensor thresholds for the Gross Range Test."
  )
```

### User Thresholds

User thresholds were calculated separately for each county due to expected and observed spatial differences in temperature (@fig-temp-mean-sd).

```{r}
#| label: fig-temp-mean-sd
#| fig-height: 8
#| fig-cap: The mean and standard deviation of temperature in each county.

p <- dat_filt %>% 
  filter(group == "county") %>% 
  plot_mean_sd_county()

ggplotly(p, tooltip = "text")

```

Most counties have substantially different number of observations for different seasons (@fig-temp-n-obs-county-month). Because temperature has a clear seasonal cycle (@fig-temp-climatology), this can weight the average towards months with more observations. To give each month equal weight, the user thresholds were based on the monthly climatology[^3].

[^3]: it was assumed that the different number of days per month was negligible

All temperature observations were first grouped by calendar month, and the average temperature for each month was calculated. The $avg_{Temp}$ was the average of these monthly averages, and $stdev_{Temp}$ was the standard deviation of the monthly averages (@eq-temp-avg, @eq-temp-stdev). With this approach, the mean for a given month will be weighted towards years with more observations than other years for that month. This is not expected to have a substantial influence on the calculated thresholds, but future iterations of this exercise could further standardize the data to account for this.

$$
avg_{temp} = sum(avg_{Jan} + avg_{Feb} + ... avg_{Dec}) / 12
$$ {#eq-temp-avg}

$$
stdev_{temp} = sd(avg_{Jan}, avg_{Feb}, ... avg_{Dec})
$$ {#eq-temp-stdev}

County statistics and user thresholds are presented in Table 2.

<br>

```{r}
# calculate thresholds based on the climatology
user_thresh <- dat_filt %>% 
  filter(group == "county_month") %>% 
  rename(mean_month = mean) %>% 
  group_by(county) %>%
  summarise(
    mean = round(mean(mean_month), digits = 3),
    stdev = round(sd(mean_month), digits = 3)
  ) %>% 
  mutate(
    qc_test = "grossrange",
    variable = "temperature_degree_c",
    user_min = mean - 3 * stdev, 
    user_max = mean + 3 * stdev
  ) 

user_thresh %>% 
  select(county, mean, stdev, user_min, user_max) %>% 
  datatable(
    options = dt_options, extensions = 'Buttons', 
    rownames = FALSE,
    caption = "Table 2: Temperature statistics and user thresholds for the Gross Range Test."
  )
```

<br>

The quality[^2] of these user thresholds may vary by county, depending on the number and distribution (in space and time) of observations. For example, there are relatively few observations for some counties compared to others (@fig-temp-n-obs). Cape Breton has the fewest observations at ~30,000 over 3 years and two stations, while Guysborough has the most at nearly 1 million over 7 years and 35 stations (@fig-temp-station-locations). Colchester and Queens counties only have one station each (@fig-temp-station-locations). For consistency, the counties with fewer observations were not pooled with counties with more observations. The user thresholds should be re-evaluated when more observations are collected.

[^2]: e.g., how representative the thresholds are of "normal" conditions through the water column and county

```{r}
#| label: fig-temp-n-obs
#| fig-height: 8
#| fig-cap: The number of dissolved oxygen observations in each county. (Note these numbers are not the same as on the CMAR measurements page because some outlier values and stations were omitted from the thresholds analysis.)

p <- dat_filt %>%
  filter(group == "county") %>% 
  plot_n_obs_county()

ggplotly(p)

```

Calculating thresholds at the county scale provides relatively coarse threshold values. Ideally, these would be resolved by depth and a smaller spatial scale (e.g., waterbody or station). However, calculating thresholds for each county and depth provides its own challenges. The data become very patchy when grouped by county, depth, and month (e.g., 169 month-county-depth combinations with 0 observations). Additionally, the same depth can represent a different part of the water column for different stations. At the Barren Island station in Guysborough county, the 15 m sensor is near the bottom. In contrast, 15 m is in the top 20 % of the water column at Tickle Island, also in Guysborough county. Finally, aggregating thresholds by county and depth would result in 141 user-defined temperature thresholds, which is more than the 2.5 person Data Governance team can reasonably manage.

Because depth was not accounted for, it is expected that observations from very shallow sensors (e.g., <= 2 m) will be assigned the [Suspect/Of Interest]{style="color: #EDA247;"} flag, despite appearing reasonable in the context of the deployment. In this case, the flag should be interpreted as ["Of Interest"]{style="color: #EDA247;"}, for highlighting a relatively warm observation.

The $user_{min}$ threshold is typically << 0 degrees Celsius (Table 2), and is therefore not expected to flag any observations. For most counties (all except Annapolis, Queens, Shelburne, and Digby), the $user_{min}$ is less than the $sensor_{min}$ for the aquameasure and vr2ar sensors. In this case, any observations less than the $sensor_{min}$ would [fail]{style="color: #DB4325;"} the Gross Range Test (i.e., the $user_{min}$ would be ignored). It may be useful for other users to apply their own $user_{min}$ threshold to highlight cold observations that are [Suspect/Of Interest]{style="color: #EDA247;"}. For example, those interested in salmonid aquaculture may wish to flag observations at or near the superchill threshold (-0.7 degree C).

## Climatological Test

The season thresholds were calculated separately for each county due to expected and observed spatial differences in temperature (@fig-temp-climatology).


::: panel-tabset

### A
```{r}
#| label: fig-temp-climatology
#| fig-height: 10
#| fig-cap: The monthly mean and standard deviation of temperature in each county.

county_month <- dat_filt %>% 
  filter(group == "county_month") %>% 
  mutate(month = month(month, label = TRUE)) %>% 
  select(-c(units, group, year, station, depth))

p <- plot_mean_sd_season(county_month, facet_county = TRUE) +
  theme(panel.spacing.y = unit(15, "lines"))

ggplotly(p)

```

### B

```{r}
#| fig-height: 8
#| fig-cap: The monthly mean and standard deviation of temperature in each county.

p <- plot_mean_sd_season_ribbon(county_month, group = "county", facet_county = FALSE)

ggplotly(p, tooltip = "text")
```
:::

Seasonal statistics and thresholds are presented in Table 3. The quality[^2] of these thresholds may vary by county and month, depending on the number and distribution (in space and time) of observations. There were no county-month groups with zero observations, although some groups had relatively few observations for many or all months (@fig-temp-n-obs-county-month; e.g., Cape Breton).

<br>

```{r}
clim_thresholds <- county_month %>% 
  mutate(
    season_min = round((mean - 3 * stdev), digits = 3),
    season_max = round((mean + 3 * stdev), digits = 3)
  ) %>% 
  select(county, month, mean, stdev, season_min, season_max) %>% 
  arrange(county, month)  

datatable(
  clim_thresholds, extensions = 'Buttons', 
  options = dt_options, rownames = FALSE,
  caption = "Table 3: Seasonal statistics and thresholds by county and month."
)
```

<br>

### Spatial & Temporal Resolution

The length of the time series varied among counties, from approximately 1 year (e.g., Cape Breton, Colchester) to over 7 years (select stations in Guysborough, Richmond, and Yarmouth counties; @fig-temp-n-years). This means that the monthly average for some county-month groups is based on only 1 year of data. 

The number of stations varied by county, from a single station each in Colchester and Queens counties, to `r nrow(dat_filt[which(dat_filt$group == "all_station" & dat_filt$county == "Guysborough"),"station"])` stations in Guysborough (@fig-temp-station-locations).

The seasonal thresholds for counties with limited years, number of observations, and/or number of stations are likely not representative of inter-annual or spatial variability in the county; however, they are adequate for this quality control exercise. These thresholds *are* representative of normal conditions of the observed deployments, and so outlying values will be flagged. If new deployments are added to these counties, it is recommended that the seasonal thresholds be re-evaluated and re-calculated if necessary. It is strongly recommended that all seasonal thresholds be re-calculated after several more years of data have been collected through the Coastal Monitoring Program.


::: panel-tabset

### Figure 5
```{r}
#| label: fig-temp-n-obs-county-month
#| fig-height: 12
#| fig-cap: The number of temperature observations in each month for each county.

 p <- plot_n_obs_month(county_month, text_size = 12) +
  theme(panel.spacing.y = unit(15, "lines"))

ggplotly(p)

```

### Figure 6
```{r}
#| label: fig-temp-n-years
#| fig-height: 12
#| fig-cap: The number of years of temperature data for each month and county.
#| message: false

p <- dat_filt %>%
  filter(group == "county_month_year") %>%
  group_by(county, month) %>%
  summarise(n = n()) %>%
  mutate(month = month(month, label = TRUE)) %>% 
  ungroup() %>% 
  plot_n_obs_month(text_size = 12, fill = "#CCCCFF") +
  scale_x_continuous("Number of Years", expand = expansion(mult = c(0, 0.1))) +
  theme(panel.spacing.y = unit(15, "lines"))

ggplotly(p)

```
:::


### Depth & Stratification

Like the user thresholds, the seasonal thresholds were not resolved by depth. This means that there was high standard deviation for counties with seasonal stratification, particularly in the summer months when stratification is typically the strongest (@fig-temp-climatology; Table 3).

The standard deviations for July through October in Inverness are the four highest overall standard deviations (Table 3). This high variability was driven by three deployments in Whycocomagh Basin in the Bras D'Or Lakes:

-   Deep Basin (May to September 2018)
-   0814x East (September to December 2020)
-   0814x West (September to December 2020)

These stations had sensors deployed above and below the thermocline. Temperatures below the thermocline, near the bottom, were typically very cold (about zero degrees for the whole Deep Basin deployment). Temperatures above the thermocline, closer to the surface, were typically much warmer, up to 25 °C in the summer (**link to Inverness report here**). The high standard deviation results in a very wide range of temperature values that would be flagged [Pass]{style="color: #009200;"}: e.g., from `r clim_thresholds[which(clim_thresholds$county == "Inverness" & clim_thresholds$month == "Aug"), "season_min"]$season_min` °C to `r clim_thresholds[which(clim_thresholds$county == "Inverness" & clim_thresholds$month == "Aug"), "season_max"]$season_max` °C in August.

The temperatures below the thermocline could be considered anomalous and removed prior to calculating the thresholds. These temperatures would then be flagged [Of Interest]{style="color: #EDA247;"}. However, the temperatures were included in the current thresholds analysis for several reasons:

1.  Together, these three deployments represent 68 % of the Inverness temperature observations (those from below the thermocline represents 30 % of the county observations).
2.  Consistency with the threshold calculation for other counties (e.g., depth was not accounted for).
3.  It is not [suspicious]{style="color: #EDA247;"} for these temperatures to be so cold in this region of Whycocomagh Basin.

There are stations withe notable depth stratification in other counties, including Guysborough, Halifax, and Lunenburg. This is reflected in the relatively high standard deviation in the summer months for these counties (@fig-temp-climatology; Table 3). Future iterations of this threshold analysis could consider resolving the seasonal thresholds by depth; however it is beyond the scope of the current exercise.

### User vs. Seasonal Thresholds


```{r}
#| label: fig-climatology-thresh
#| fig-cap: Monthly dissolved oxygen mean +/- 1 standard deviation. Solid orange lines indicate the climatology thresholds; dashed organge lines indicate the grossrange user thresholds.

# user_col <- "#EDA247"
# season_col <- 1
# 
# x <- county_month %>% 
#   left_join(
#     user_thresh %>% 
#       select(-c(variable, qc_test, mean, stdev)), 
#     by = "county") %>%
#   left_join()
# 
# plot_mean_sd_season(x, facet_county = TRUE) +
#   geom_ribbon(
#     data = x,
#     aes(month, 
#         ymin = mean - 3 * stdev,
#         ymax = mean + 3 * stdev)
#   )
# 
#     geom_hline(data = x, aes(yintercept = user_min), col = 2)
#   
#   
#   
# ggplot(x, aes(month, mean)) +
#     geom_point(size = 1) +
#     geom_errorbar(
#       aes(ymin = mean - stdev, ymax = mean + stdev), width = 0
#     ) +
#     scale_x_discrete(
#       name = "", breaks = c("Jan", "Mar", "May", "Jul", "Sep", "Nov")
#     ) +
#   facet_wrap(~county, ncol = 3) +
#   geom_hline(data = x, aes(yintercept = user_min))

  
```











